Synthesizing Symmetric Lenses

### 2018-07-04 papers/SymmetricLenses/main.tex
Finished proof of Theorem 7
Finished like a third of Theorem 3

Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$. Then
\begin{align*}
H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
\end{align*}
\end{enumerate}
\end{proof}
\end{document}

>>>>

Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$. Then
\begin{align*}
H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
\end{align*}
\end{enumerate}
\end{proof}
\begin{theorem*}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{center}
    \begin{tabular}{rcl}
      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
    \end{tabular}
  \end{center}
\end{theorem*}
\begin{proof}
Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
\begin{align*}
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
\end{align*}
Define the simple symmetric lens $\lceil \ell \rceil$ by 
\begin{align*}
\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
\end{align*}
We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
\begin{align*}
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
\end{align*}
\end{proof}

\begin{theorem*}
  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
  \begin{enumerate}
  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
      \Lens.\PutLOf{t}{s} = s}$
  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
      \Lens.\PutROf{s}{t} = t}$
  \end{enumerate}
\end{theorem*}
\begin{proof}
Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutLOf{t}{s} = s\}$. We want to show that
$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}H(V_{\ell}(s)) = \mathbb{H}^{\leftarrow}(S \; | \; \ell, T)$$
We proceed by induction over $\ell$. 
\begin{center}
  \begin{tabular}{l@{\ }l@{\ }l@{\ }l}
            & & \GBar{} & $\ConcatLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\SwapLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\OrLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\ComposeLensOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\MergeROf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\MergeLOf{\Lens_1}{\Lens_2}$ \\
            & & \GBar{} & $\InvertOf{\Lens}$ \\
  \end{tabular}
\end{center}
\begin{enumerate}
\item
($\ell = \IdentityLensOf{S}$)

Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\PutLOf{t}{s} = s\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.

Consequently, $H(V_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\leftarrow}(S \; | \; \IdentityLensOf{S}, T)$$
\item
($\ell = \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$)

Observe that 
$$V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutLOf{t}{s} = s\} = \{t \in \mathcal{L}(T) \; | \; s = s\} = \mathcal{L}(T)$$
Consequently
$$\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) = \sum_{s \in \mathcal{L}(S)}P_{\mathcal{S}}(s)H(\mathcal{L}(T)) = H(\mathcal{L}(T)) = \mathbb{H}(T) = \mathbb{H}^{\leftarrow}(S \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, T)$$
\item
($\ell = \IterateLensOf{\Lens} : S^{*p} \Leftrightarrow T^{*q}$)

Observe that 
\begin{align*}
V_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutLOf{(t_1 \cdot \ldots \cdot t_m)}{(s_1 \cdot \ldots \cdot s_n)} = s_1 \cdot \ldots \cdot s_n\}\\
&= \{t_1 \cdot \ldots \cdot t_n \; | \; \ell.\PutLOf{t_i}{s_i} = s_i \text{ for }1 \leq i \leq n\}\\
&= V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n)
\end{align*}
Consequently
\begin{align*}
\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n))\\
&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)\\
&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)
\end{align*}
We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j))$:
\begin{align*}
\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V_{\ell}(s_j))\\
&= \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))
\end{align*}
Consequently,
\begin{align*}
\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \mathbb{H}^{\leftarrow}(S \; | \; \ell, T)\\
&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
&= \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)
\end{align*}
Unrolling the term 
$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)$$ 
$(n-1)$ more times gives
$$\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
= \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) * (1-p)\sum_n n * p^n  = \frac{p}{1-p} * \mathbb{H}^{\leftarrow}(S \; | \; \ell, T)$$
which is what we wanted to show.
\end{enumerate}

\end{proof}
\end{document}
